### 2021-03-24 paper/header.tex
First version of Proceeding.

% **************GENERATED FILE, DO NOT EDIT**************

\title{Running faster scripts with DaemonMode.jl}

\author[1]{Daniel Molina}
\affil[1]{DaSCI Andalusian Institute of Data Science and Computational
  Intelligence, Department of Computer Science, University of Granada, Spain}

\keywords{Julia, Optimization, Tool, Script, Time-to-first plot problem}

\hypersetup{
  pdftitle = {Running faster scripts with DaemonMode.jl, JuliaCon proceeding},
pdfsubject = {JuliaCon 2021 Proceedings},
pdfauthor = {Daniel Molina},
pdfkeywords = {Julia, Optimization, Tool, Scripts, Time-to-first plot problem},
}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: