### 2021-10-24 paper/header.tex
Add affiliation.

% **************GENERATED FILE, DO NOT EDIT**************

\title{Running faster scripts with DaemonMode.jl}

\author[1, 2]{Daniel Molina}
\affil[1]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, Spain}
\affil[2]{Department of Computer Science, University of Granada, Spain}

\keywords{Julia, Optimization, Tool, Script, Time-to-first plot}

\hypersetup{
pdftitle = {Running faster scripts with DaemonMode.jl},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Daniel Molina},
pdfkeywords = {Julia, Optimization, Tool, Script, Time-to-first plot},
}